---
title: "Estimations_robsurvey"
author: "Loan Rottier, Simon Simus and Maelys Rapenne"
date: "2025-11-18"
output: html_document
---

```{r}
library(readr) 
library(tidyverse) 
library(ggplot2) 
library(sampling) 
library(survey)
library(robsurvey)

```

```{r}
data(MU284)
```

```{r}
Q1 <- quantile(MU284$P85, 0.25)
Q3  <- quantile(MU284$P85, 0.75) 
IQR <- Q3 - Q1
lower <- Q1 - 1.5 * IQR 
upper <- Q3 + 1.5 * IQR
outliers <- MU284$LABEL[MU284$P85 < lower | MU284$P85 > upper]
MU284$outlier <- ifelse(seq_len(nrow(MU284)) %in% outliers, 1, 0)
```

# Winsorization for the mean in SRSWOR
```{r}

set.seed(1)
N <- nrow(MU284)
n <- 70

ech_index <- srswor(n, N)
echantillon_srswor <- MU284[ech_index == 1, ]

#create a vector of size n with weight N/n for all element
poids_srswor <- rep(N/n, n)
#create the survey design with weight 
ech.srswor <- svydesign(id=~LABEL, weights=poids_srswor, fpc=rep(N, n), data=echantillon_srswor)
```


```{r}
#real mean : 
P85_mean_pop <- mean(MU284$P85)
P85_mean_pop
#Horvitz thomson estimator for the mean (with standart deviation)
estimate_mean <- svymean(~P85, ech.srswor)
estimate_mean
```
## Once-Winsorized mean

There are 2 or more to compute the once-winsorized mean:
Using the robsurvey package function svymean_k_winsorized with k = 1
We use the 2nd biggest value as the cutoff then we cut the value of the biggest value to the cutoff (the 2nd biggest value)

or we can write the formula of the once-winsorized mean:
y_bar(1) = y_bar - (y_(n) - y_(n-1)) / n
where y_(n) is the biggest value of the sample and y_(n-1) the 2nd biggest value of the sample


```{r}
P85_sample <- echantillon_srswor$P85
y_bar <- mean(P85_sample)
P85_sorted <- sort(P85_sample, decreasing = TRUE)
y_n <- P85_sorted[1]       # y_(n)
y_n_moins_1 <- P85_sorted[2] # y_(n-1)

y_bar_1_manuel <- y_bar - (y_n - y_n_moins_1) / n
y_bar_1_manuel
```


Winsorization by the k biggest values

Replace the k biggest values of our sample (SRSWOR) by the k+1 biggest value in the sample 

Try the Rivest winsorization with k = 1 by taking the takeoff as the second biggest values
```{r}
svymean_1_winsorized <- svymean_k_winsorized(~P85, 
                                              ech.srswor, k = 1)
svymean_1_winsorized
#MSE
mse(svymean_1_winsorized)
```

We try to find the cutoff that minimize the MSE

In order to compute the MSE 


With the Winsorization by Rivest, it's with k=1 but we can try some other value of k and see the impact on estimate mean


```{r}
means_k_winso <- numeric(15)
for (k in 1:15){
  means_k_winso[k] <- coef(svymean_k_winsorized(~P85, ech.stsi, k = k))
}
plot(means_k_winso, ylim = c(23,30), xlab = "Value of k", ylab = "mean")
abline(h = P85_mean, col = "red", lwd = 2)
```

```{r}
means_k_winso <- numeric(15)
for (k in 1:15){
  means_k_winso[k] <- SE(svymean_k_winsorized(~P85, ech.stsi, k = k))
}
plot(means_k_winso, xlab = "Value of k", ylab = "SE of the estimated means")
abline(h = P85_mean, col = "red", lwd = 2)
```

##Winsorization with the Searl's approach : 

True mean

```{r}
P85_mean <- mean(MU284$P85)
P85_mean
```


Documentation

Characteristic. Population mean or total. Let μ denote the estimated winsorized population mean; then, the estimated winsorized total is given by \$ \\hat{N} \\mu \$ with \$\\hat{N} = \\sum w_i\$ , where summation is over all observations in the sample.

Modes of winsorization. The amount of winsorization can be specified in relative or absolute terms:

-   Relative: By specifying LB and UB, the method winsorizes the LB · 100% of the smallest observations and the (1 - UB) · 100% of the largest observations from the data.

-   Absolute: By specifying argument k in the functions with the "infix" *k* in their name (e.g., svymean_k_winsorized), the largest k observations are winsorized, 0 \< k \< n, where n denotes the sample size. E.g., k = 2 implies that the largest and the second largest observation are winsorized. Variance estimation. Large-sample approximation based on the influence function; see Huber and Ronchetti (2009, Chap. 3.3) and Shao (1994). Two estimators are available:

    -   simple_var = FALSE Variance estimator of the winsorized mean/ total. The estimator de- pends on the estimated probability density function evaluated at the winsorization thresh- olds, which can be – depending on the context – numerically unstable. As a remedy, a simplified variance estimator is available by setting simple_var = TRUE.

    -   simple_var = TRUE Variance is approximated using the variance estimator of the trimmed mean/ total.


Optimal R

R solution of \$\\frac{R - \\mu}{n-1} = \\mathbb{E}\_F(max(Y-R,0))\$

On estime l'espérance puis on résoud \$H(R) = \\frac{R - \\mu}{n-1} - {1\\over n} \\sum\_{i=1}\^n max(Y-R,0)) = 0\$

```{r}
mu_pop <- mean(MU284$P85)
n <- length(stsrswor_data)
H_R <- function(R, mu, n, sample){
  terme1 <- (R - mu) / (n - 1)
  terme2 <- sum(pmax(sample - R, 0))/n
  return(terme1 - terme2)
}
R_esp_max <- uniroot(f = H_R, interval = c(min(MU284$P85), max(MU284$P85)),
                     mu = mu_pop, n = n, sample = stsrswor_data$P85)$root
R_esp_max
```

find R

```{r}
MU284 %>% arrange(desc(P85)) %>% select(P85)
```

2 are superior than R so k = 2

However the ratio R depend about the empirical distribution and then on the data (and possible outliers)


Then we will try the approach of Searl. The cutoff is based on a quantile. The difficulty is to find the best quantile. We will try some quantile and keep the quantile with the lowest MSE. 

```{r}
quantiles <- seq(0.9, 0.99, by = 0.01)
mse_values <- numeric(length(quantiles))
for (i in seq_along(quantiles)) {
  UB <- quantiles[i]
  winsorized_mean <- svymean_winsorized(~P85, ech.srswor, LB = 0, UB = UB)
  mse_values[i] <- mse(winsorized_mean)
}
mse_values
best_quantile <- quantiles[which.min(mse_values)]
best_quantile
```
The best quantile seem to be the 0.98. 


### Winsorization with MonteCarlo
```{r}
nsimu <- 1000
estimations_HT <- numeric(nsimu)
estimations_k1 <- numeric(nsimu) 
estimations_UB_fixed <- numeric(nsimu)
```
For Searl's method:
```{r}

for (i in 1:nsimu) {
  
  ech_index <- srswor(n, N)
  sample_si <- MU284[ech_index == 1, ]
  
  weight_si <- rep(N/n, n)
  ech.si <- svydesign(id=~LABEL, weights=weight_si, fpc=rep(N, n), data=sample_si)
  #for comparison with H-T estimator
  mean_HT <- svymean(~P85, ech.si)
  estimations_HT[i] <- coef(mean_HT)
  
  # winsorized mean with k = 1
  mean_k1 <- svymean_k_winsorized(~P85, ech.si, k = 1)
  estimations_k1[i] <- coef(mean_k1) 
  
  # Compute the mean with searl's fixed UB 
  mean_UB_fixed <- svymean_winsorized(~P85, ech.si, LB = 0, UB = 0.97)
  estimations_UB_fixed[i] <- coef(mean_UB_fixed) 
  
}


```
then compute the bias, variance and MSE for both methods
```{r}
cat("with rivest method (k=1)")

bias_k1 <- mean(estimations_k1) - P85_mean_pop
var_k1 <- var(estimations_k1)
mse_k1 <- bias_k1^2 + var_k1
bias_k1
var_k1
mse_k1

cat("with searl method (fixed UB)")

bias_UB_fixed <- mean(estimations_UB_fixed) - P85_mean_pop
var_UB_fixed <- var(estimations_UB_fixed)
mse_UB_fixed <- bias_UB_fixed^2 + var_UB_fixed
bias_UB_fixed
var_UB_fixed
mse_UB_fixed

cat("comparison with Horvitz-Thomson estimator")
bias_HT <- mean(estimations_HT) - P85_mean_pop
var_HT <- var(estimations_HT)
mse_HT <- bias_HT^2 + var_HT
bias_HT
var_HT
mse_HT
```
The H-T have a small biais because we don't 
we can see that with rivest's and searl's method we have a bias
but the variance and mse are lower than the H-T estimator



# Create Stratum

```{r}
MU284$strate <- cut(MU284$P75, breaks = quantile(MU284$P75, probs = c(0, 1/2, 28.5/30, 1)), include.lowest = TRUE, labels = c("basse", "moyenne", "haute"))
MU284 <- MU284 %>% mutate(strate_id = case_when(
  strate == "basse" ~ 1,
  strate == "moyenne" ~ 2,
  strate == "haute" ~ 3 ))
```

```{r}
ggplot(MU284, aes(x = P75, y = P85, color = factor(strate))) + 
  geom_point() + 
  labs( title = "Scatter plot entre P75 et P85", x = "Population en 1975 (P75)", y = "Population en 1985 (P85)" ) + 
  theme_minimal()
```

## Compute the neyman allocation

```{r}
N_H <- table(MU284$strate)
n <- 70 
table(MU284$strate) 
std <- c() 
for (i in 1:3){ 
  std[i] <- sqrt(var(MU284$P85[MU284$strate_id == i])) 
  } 
tab <- table(MU284$strate)*std 
sum_NL <- sum(tab) 
n_H <- n * tab / sum_NL 
round(n_H) #pas ok car n_3 = 34 > N_3 = 15
```

we restant with fixing $n_3$.

```{r}
n_3 <- table(MU284$strate)['haute'] 
n_12 <- n - n_3
tab <- table(MU284$strate)[1:2] * std[1:2] 
sum_NL <- sum(tab) 
n_H <- n_12 * tab / sum_NL 
n_H['haute'] <- n_3 
n_H <- round(n_H)
n_H
```

```{r}
#MU284_sorted <- MU284 %>% arrange(strate)
#sizes <- c(11, 44, 15)
#stsrswor = strata(MU284_sorted,"strate",size=sizes,method="srswor")
#stsrswor_data = getdata(MU284,stsrswor)
```

```{r}
#creae sample of size nh for each stratum
ech1 <- sample(1:nrow(MU284[MU284$strate=="basse",]), size = n_H["basse"])
ech2 <- sample(1:nrow(MU284[MU284$strate=="moyenne",]), size = n_H["moyenne"])
ech3 <- sample(1:nrow(MU284[MU284$strate=="haute",]), size = n_H["haute"])
sample1 <- MU284[MU284$strate=="basse",]
sample1 <- sample1[ech1,]
sample2 <- MU284[MU284$strate=="moyenne",]
sample2 <- sample2[ech2,]
sample3 <- MU284[MU284$strate=="haute",]
sample3 <- sample3[ech3,]
stsrswor_data <- rbind(sample1, sample2, sample3) 
```

```{r}
poids_stsi=c(rep(N_H[1]/n_H[1],n_H[1]),rep(N_H[2]/n_H[2],n_H[2]),rep(N_H[3]/n_H[3],n_H[3])) 
ech.stsi=svydesign(id=~LABEL,strata=~strate,weights=poids_stsi, fpc=c(rep(N_H[1], n_H[1]), rep(N_H[2], n_H[2]), rep(N_H[3], n_H[3])), data=stsrswor_data) 
```

#  Winsorization for the total
## Winsorization in SRSWOR: 

```{r}

nsimu <- 1000

data(MU284)
N <- nrow(MU284)
n <- 70

true_total <- sum(MU284$P85)

results_ht <- numeric(nsimu)
results_k1 <- numeric(nsimu)
results_ub <- numeric(nsimu)

for (i in 1:nsimu) {
  
  s <- srswor(n, N)
  sample_data <- MU284[s == 1, ]
  
  sample_data$weights <- N / n
  sample_data$fpc <- N
  
  design_obj <- svydesign(
    id = ~1,
    weights = ~weights,
    fpc = ~fpc,
    data = sample_data
  )
  
  results_ht[i] <- coef(svytotal(~P85, design_obj))
  results_k1[i] <- coef(svytotal_k_winsorized(~P85, design_obj, k = 1))
  results_ub[i] <- coef(svytotal_winsorized(~P85, design_obj, LB = 0, UB = 0.97))
}

mse_ht <- mean((results_ht - true_total)^2)
mse_k1 <- mean((results_k1 - true_total)^2)
mse_ub <- mean((results_ub - true_total)^2)

var_ht <- var(results_ht)
var_k1 <- var(results_k1)
var_ub <- var(results_ub)

bias_ht <- mean(results_ht) - true_total
bias_k1 <- mean(results_k1) - true_total
bias_ub <- mean(results_ub) - true_total

final_results <- data.frame(
  Estimator = c("Horvitz-Thompson", "Winsorized (k=1)", "Winsorized (UB=0.95)"),
  MSE = c(mse_ht, mse_k1, mse_ub),
  Variance = c(var_ht, var_k1, var_ub),
  Bias = c(bias_ht, bias_k1, bias_ub)
)

print(final_results)

```


## Winsorization with stratum for STSRSWOR : 
we see before that with the neyman allocation,
we find n1 =11, n2 = 44, n3 = 15
since in the last statrum we have N3 = 15, we can't do winsorization 
 
So we will realize stratification for stratum 1 and 2
Monte carlo : 



```{r}
nsimu <- 1000 
estimations_HT_strat <- numeric(nsimu)
estimations_k1_strat <- numeric(nsimu)
estimations_UB_fixed_strat <- numeric(nsimu)
for (i in 1:nsimu) {
  sample_si_list <- list()
  
  data_basse <- MU284[MU284$strate == "basse", ]
  N_basse <- nrow(data_basse)
  n_basse <- n_H["basse"]
  ech_index_basse <- srswor(n_basse, N_basse)
  sample_basse <- data_basse[ech_index_basse == 1, ]
  
  data_moyenne <- MU284[MU284$strate == "moyenne", ]
  N_moyenne <- nrow(data_moyenne)
  n_moyenne <- n_H["moyenne"]
  ech_index_moyenne <- srswor(n_moyenne, N_moyenne)
  sample_moyenne <- data_moyenne[ech_index_moyenne == 1, ]
  
  #take all values in strat haute
  data_haute <- MU284[MU284$strate == "haute", ]
  
  
  sample_si <-rbind(sample_basse, sample_moyenne, data_haute)
  weight_si <- c(rep(N_basse/n_basse, n_basse), rep(N_moyenne/n_moyenne, n_moyenne), rep(1, n_H["haute"]))
  ech.si <- svydesign(id=~LABEL, strata=~strate, weights=weight_si, fpc=c(rep(N_basse, n_basse), rep(N_moyenne, n_moyenne), rep(N_H["haute"], n_H["haute"])), data=sample_si)
  
  #for comparison with H-T estimator
  res_ht <- svyby(~P85, ~strate, ech.si, svytotal)
  estimations_HT_strat[i] <- sum(res_ht$P85)
  

  # winsorized total with k = 1 in each stratum
  res_k1 <- svyby(~P85, ~strate, ech.si, svytotal_k_winsorized, k = 5)
  total_k1_mixte <- res_k1$P85[res_k1$strate == "basse"] + 
                    res_k1$P85[res_k1$strate == "moyenne"] + 
                    res_ht$P85[res_ht$strate == "haute"]
  
  estimations_k1_strat[i] <- total_k1_mixte
  
  # Compute the total with searl's fixed UB
  res_ub <- svyby(~P85, ~strate, ech.si, svytotal_winsorized, LB = 0, UB = 0.8)
  total_ub_mixte <- res_ub$P85[res_ub$strate == "basse"] + 
                    res_ub$P85[res_ub$strate == "moyenne"] + 
                    res_ht$P85[res_ht$strate == "haute"]
  estimations_UB_fixed_strat[i] <- total_ub_mixte
}

```


```{r}

Y_total <- sum(MU284$P85)
Y_HT <- mean(estimations_HT_strat)
Y_UB_fixed_total <- mean(estimations_UB_fixed_strat)
Y_searl_total <- mean(estimations_k1_strat)
  
Y_total
Y_HT
Y_UB_fixed_total
Y_searl_total 

cat("compare mse")
mse_ht <- mean((estimations_HT_strat - Y_total)^2)
mse_k1 <- mean((estimations_k1_strat - Y_total)^2)
mse_ub <- mean((estimations_UB_fixed_strat - Y_total)^2)


mse_ht
mse_k1
mse_ub

cat("compare bias")
bias_ht <- mean(estimations_HT_strat) - Y_total
bias_k1 <- mean(estimations_k1_strat) - Y_total
bias_ub <- mean(estimations_UB_fixed_strat) - Y_total

bias_ht
bias_k1
bias_ub

cat("compare variance")
var_ht <- var(estimations_HT_strat)
var_k1 <- var(estimations_k1_strat)
var_ub <- var(estimations_UB_fixed_strat)

var_ht
var_k1
var_ub

```
We can remark that the variance and mse are higher with the winsorization methods than with the H-T estimator.

It's due tu the fact that we dont have outlier in our sample. So we will add some outlier in the dataset in order to see the effect of winsorization in presence of outlier.

creation of a new dataset with outlier
## Creation of a new dataset with outliers
```{r}
MU284_OUTLIERS <- MU284

indices_basse <- which(MU284_OUTLIERS$strate == "basse")

#change the first 2 indices in strat basse and multiply it by 4 and 8
MU284_OUTLIERS$P85[indices_basse[1]] <- MU284_OUTLIERS$P85[indices_basse[1]] * 4
MU284_OUTLIERS$P85[indices_basse[2]] <- MU284_OUTLIERS$P85[indices_basse[2]] * 8


indices_moyenne <- which(MU284_OUTLIERS$strate == "moyenne")
#change the first 2 indices in strat moyenne and multiply it by 4 and 8
MU284_OUTLIERS$P85[indices_moyenne[1]] <- MU284_OUTLIERS$P85[indices_moyenne[1]] * 5
MU284_OUTLIERS$P85[indices_moyenne[2]] <- MU284_OUTLIERS$P85[indices_moyenne[2]] * 4


print("see the data with outliers in comparaison without outliers")
by(MU284_OUTLIERS$P85, MU284_OUTLIERS$strate, summary)
```
```{r}
set.seed(2)
nsimu <- 1000 
estimations_HT_strat <- numeric(nsimu)
estimations_k1_strat <- numeric(nsimu)
estimations_UB_fixed_strat <- numeric(nsimu)
for (i in 1:nsimu) {
  sample_si_list <- list()
  
  data_basse <- MU284_OUTLIERS[MU284_OUTLIERS$strate == "basse", ]
  N_basse <- nrow(data_basse)
  n_basse <- n_H["basse"]
  ech_index_basse <- srswor(n_basse, N_basse)
  sample_basse <- data_basse[ech_index_basse == 1, ]
  
  data_moyenne <- MU284_OUTLIERS[MU284_OUTLIERS$strate == "moyenne", ]
  N_moyenne <- nrow(data_moyenne)
  n_moyenne <- n_H["moyenne"]
  ech_index_moyenne <- srswor(n_moyenne, N_moyenne)
  sample_moyenne <- data_moyenne[ech_index_moyenne == 1, ]
  
  #take all values in strat haute
  data_haute <- MU284_OUTLIERS[MU284_OUTLIERS$strate == "haute", ]
  
  
  sample_si <-rbind(sample_basse, sample_moyenne, data_haute)
  weight_si <- c(rep(N_basse/n_basse, n_basse), rep(N_moyenne/n_moyenne, n_moyenne), rep(1, n_H["haute"]))
  ech.si <- svydesign(id=~LABEL, strata=~strate, weights=weight_si, fpc=c(rep(N_basse, n_basse), rep(N_moyenne, n_moyenne), rep(N_H["haute"], n_H["haute"])), data=sample_si)
  
  #for comparison with H-T estimator
  res_ht <- svyby(~P85, ~strate, ech.si, svytotal)
  estimations_HT_strat[i] <- sum(res_ht$P85)
  

  # winsorized total with k = 1 in each stratum
  res_k1 <- svyby(~P85, ~strate, ech.si, svytotal_k_winsorized, k = 1)
  total_k1_mixte <- res_k1$P85[res_k1$strate == "basse"] + 
                    res_k1$P85[res_k1$strate == "moyenne"] + 
                    res_ht$P85[res_ht$strate == "haute"]
  
  estimations_k1_strat[i] <- total_k1_mixte
  
  # Compute the total with searl's fixed UB
  res_ub <- svyby(~P85, ~strate, ech.si, svytotal_winsorized, LB = 0, UB = 0.9)
  total_ub_mixte <- res_ub$P85[res_ub$strate == "basse"] + 
                    res_ub$P85[res_ub$strate == "moyenne"] + 
                    res_ht$P85[res_ht$strate == "haute"]
  estimations_UB_fixed_strat[i] <- total_ub_mixte
}
```


```{r}

Y_total <- sum(MU284_OUTLIERS$P85)
Y_HT <- mean(estimations_HT_strat)
Y_UB_fixed_total <- mean(estimations_UB_fixed_strat)
Y_searl_total <- mean(estimations_k1_strat)
  
Y_total
Y_HT
Y_UB_fixed_total
Y_searl_total 



cat("compare bias")
bias_ht <- mean(estimations_HT_strat) - Y_total
bias_k1 <- mean(estimations_k1_strat) - Y_total
bias_ub <- mean(estimations_UB_fixed_strat) - Y_total

bias_ht
bias_k1
bias_ub

cat("compare variance")
var_ht <- var(estimations_HT_strat)
var_k1 <- var(estimations_k1_strat)
var_ub <- var(estimations_UB_fixed_strat)

var_ht
var_k1
var_ub

cat("compare mse")
mse_ht <- mean((estimations_HT_strat - Y_total)^2)
mse_k1 <- mean((estimations_k1_strat - Y_total)^2)
mse_ub <- mean((estimations_UB_fixed_strat - Y_total)^2)

mse_ht
mse_k1
mse_ub
```
We can see that with the presence of outliers, the winsorization methods give better results than the H-T estimator in terms of variance and MSE.
As expected, the HT-estimator has a lowest biais since the winsorization methods introduce a biais in the estimation.


















