---
title: "Estimations_robsurvey"
author: "Loan Rottier"
date: "2025-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr) 
library(tidyverse) 
library(ggplot2) 
library(sampling) 
library(survey)
library(robsurvey)
```

```{r}
data(MU284)
```

```{r}
Q1 <- quantile(MU284$P85, 0.25)
Q3  <- quantile(MU284$P85, 0.75) 
IQR <- Q3 - Q1
lower <- Q1 - 1.5 * IQR 
upper <- Q3 + 1.5 * IQR
outliers <- MU284$LABEL[MU284$P85 < lower | MU284$P85 > upper]
MU284$outlier <- ifelse(seq_len(nrow(MU284)) %in% outliers, 1, 0)
```

## Stratum

```{r}
MU284$strate <- cut(MU284$P75, breaks = quantile(MU284$P75, probs = c(0, 1/2, 28.5/30, 1)), include.lowest = TRUE, labels = c("basse", "moyenne", "haute"))
MU284 <- MU284 %>% mutate(strate_id = case_when(
  strate == "basse" ~ 1,
  strate == "moyenne" ~ 2,
  strate == "haute" ~ 3 ))
```

```{r}
ggplot(MU284, aes(x = P75, y = P85, color = factor(strate))) + 
  geom_point() + 
  labs( title = "Scatter plot entre P75 et P85", x = "Population en 1975 (P75)", y = "Population en 1985 (P85)" ) + 
  theme_minimal()
```

## Sample

```{r}
N_H <- table(MU284$strate)
# calcul n_h
n <- 70 #choix de n
table(MU284$strate) 
std <- c() 
for (i in 1:3){ 
  std[i] <- sqrt(var(MU284$P85[MU284$strate_id == i])) 
  } 
tab <- table(MU284$strate)*std 
sum_NL <- sum(tab) 
n_H <- n * tab / sum_NL 
round(n_H) #pas ok car n_3 = 34 > N_3 = 15
```

On recommence en fixant $n_3$.

```{r}
n_3 <- table(MU284$strate)['haute'] 
n_12 <- n - n_3
tab <- table(MU284$strate)[1:2] * std[1:2] 
sum_NL <- sum(tab) 
n_H <- n_12 * tab / sum_NL 
n_H['haute'] <- n_3 
n_H <- round(n_H)
n_H
```

```{r}
MU284_sorted <- MU284 %>% arrange(strate)
sizes <- c(11, 44, 15)
stsrswor = strata(MU284_sorted,"strate",size=sizes,method="srswor")
stsrswor_data = getdata(MU284,stsrswor)
```

```{r}
#parce que la ligne d'au-dessus ne marche pas
ech1 <- sample(1:nrow(MU284[MU284$strate=="basse",]), size = n_H["basse"])
ech2 <- sample(1:nrow(MU284[MU284$strate=="moyenne",]), size = n_H["moyenne"])
ech3 <- sample(1:nrow(MU284[MU284$strate=="haute",]), size = n_H["haute"])
sample1 <- MU284[MU284$strate=="basse",]
sample1 <- sample1[ech1,]
sample2 <- MU284[MU284$strate=="moyenne",]
sample2 <- sample2[ech2,]
sample3 <- MU284[MU284$strate=="haute",]
sample3 <- sample3[ech3,]
stsrswor_data <- rbind(sample1, sample2, sample3) 
```

```{r}
poids_stsi=c(rep(N_H[1]/n_H[1],n_H[1]),rep(N_H[2]/n_H[2],n_H[2]),rep(N_H[3]/n_H[3],n_H[3])) 
ech.stsi=svydesign(id=~LABEL,strata=~strate,weights=poids_stsi, fpc=c(rep(N_H[1], n_H[1]), rep(N_H[2], n_H[2]), rep(N_H[3], n_H[3])), data=stsrswor_data) 
```

## Mean

True mean

```{r}
P85_mean <- mean(MU284$P85)
P85_mean
```

### Winsorization

Winsorization with lower and upper bounds : no winsorization for the lower bound and 0.9 percentile for the lower bound

#### Documentation

Characteristic. Population mean or total. Let μ denote the estimated winsorized population mean; then, the estimated winsorized total is given by \$ \\hat{N} \\mu \$ with \$\\hat{N} = \\sum w_i\$ , where summation is over all observations in the sample.

Modes of winsorization. The amount of winsorization can be specified in relative or absolute terms:

-   Relative: By specifying LB and UB, the method winsorizes the LB · 100% of the smallest observations and the (1 - UB) · 100% of the largest observations from the data.

-   Absolute: By specifying argument k in the functions with the "infix" *k* in their name (e.g., svymean_k_winsorized), the largest k observations are winsorized, 0 \< k \< n, where n denotes the sample size. E.g., k = 2 implies that the largest and the second largest observation are winsorized. Variance estimation. Large-sample approximation based on the influence function; see Huber and Ronchetti (2009, Chap. 3.3) and Shao (1994). Two estimators are available:

    -   simple_var = FALSE Variance estimator of the winsorized mean/ total. The estimator de- pends on the estimated probability density function evaluated at the winsorization thresh- olds, which can be – depending on the context – numerically unstable. As a remedy, a simplified variance estimator is available by setting simple_var = TRUE.

    -   simple_var = TRUE Variance is approximated using the variance estimator of the trimmed mean/ total.

```{r}
svymean_winsorized(~P85, ech.stsi, LB = 0, UB = 0.90)
```

Winsorization by the k bigger values

```{r}
svymean_k_winsorized(~P85, ech.stsi, k = 5)
```

Try different values of k

```{r}
means_k_winso <- numeric(15)
for (k in 1:15){
  means_k_winso[k] <- coef(svymean_k_winsorized(~P85, ech.stsi, k = k))
}
plot(means_k_winso, ylim = c(23,30), xlab = "Value of k", ylab = "mean")
abline(h = P85_mean, col = "red", lwd = 2)
```

```{r}
means_k_winso <- numeric(15)
for (k in 1:15){
  means_k_winso[k] <- SE(svymean_k_winsorized(~P85, ech.stsi, k = k))
}
plot(means_k_winso, xlab = "Value of k", ylab = "SE of the estimated means")
abline(h = P85_mean, col = "red", lwd = 2)
```

Optimal R

R solution de \$\\frac{R - \\mu}{n-1} = \\mathbb{E}\_F(max(Y-R,0))\$

On estime l'espérance puis on résoud \$H(R) = \\frac{R - \\mu}{n-1} - {1\\over n} \\sum\_{i=1}\^n max(Y-R,0)) = 0\$

```{r}
mu_pop <- mean(MU284$P85)
n <- length(stsrswor_data)
H_R <- function(R, mu, n, sample){
  terme1 <- (R - mu) / (n - 1)
  terme2 <- sum(pmax(sample - R, 0))/n
  return(terme1 - terme2)
}
R_esp_max <- uniroot(f = H_R, interval = c(min(MU284$P85), max(MU284$P85)),
                     mu = mu_pop, n = n, sample = stsrswor_data$P85)$root
R_esp_max
```

Trouver R

```{r}
MU284 %>% arrange(desc(P85)) %>% select(P85)
```

2 sont supérieurs à R donc k = 2

### Trimmed

#### Documentation

Characteristic. Population mean or total. Let μ denote the estimated trimmed population mean; then, the estimated trimmed total is given by \$\\hat{N} \\mu\$ with \$\\hat{N} = \\sum w_i\$, where summation is over all observations in the sample.

Trimming. The methods trims the LB · 100% of the smallest observations and the (1 - UB) · 100% of the largest observations from the data. Variance estimation. Large-sample approximation based on the influence function; see Huber and Ronchetti (2009, Chap. 3.3) and Shao (1994).

#### Functions

```{r}
svymean_trimmed(~P85, ech.stsi, LB = 0, UB = 0.90)
```

### Other estimators

Weighted Huber and Tukey M-estimator of the population mean and total (robust Horvitz-Thompson estimator)

```{r}
mean_huber <- svymean_huber(~P85, ech.stsi, k = 350)
mean_huber
```

```{r}
svymean_tukey(~P85, ech.stsi, k = 350)
```

Essayer les autres paramètres :

-   k [double] robustness tuning constant (0 \< k ≤ ∞).

-   type [character] type of method: "rwm" or "rht".

-   asym [logical] if TRUE, an asymmetric Huber psi-function is used (default: FALSE).

```{r}
svymean_tukey(~P85, ech.stsi, k = 350, type = "rht")
```

```{r}
svymean_tukey(~P85, ech.stsi, k = 3, asym =  TRUE) #asym change rien ?
```

Dalen, see weighted for more details

```{r}
svymean_dalen(~P85, ech.stsi, censoring = 350)
```

```{r}
svymean_dalen(~P85, ech.stsi, censoring = 350, type = 'Z3')
```

Ratio

-   object an object of class [ratio], e.g., result of the Huber ratio M-estimator svyratio_huber.

-   total [numeric] vector of population totals of the auxiliary variables.

-   N [numeric] population size (see also N_unknown.

```{r}
svymean_ratio(mean_huber,MU284$P75, 284)
```

```{r}
coef(mean_huber)
```

## Total

True total

```{r}
P85_total <- sum(MU284$P85)
```

### Winsorization

```{r}
svytotal_winsorized(~P85, ech.stsi, LB = 0, UB = 0.90)
```

```{r}
svytotal_k_winsorized(~P85, ech.stsi, k = 2)
```

```{r}
totals_k_winso <- numeric(15)
for (k in 1:15){
  totals_k_winso[k] <- coef(svytotal_k_winsorized(~P85, ech.stsi, k = k))
}
plot(totals_k_winso, ylim = c(6600,8400), xlab = "Value of k", ylab = "SE of the total estimation")
abline(h = P85_total, col = "red", lwd = 2)
```

```{r}
totals_k_winso <- numeric(15)
for (k in 1:15){
  totals_k_winso[k] <- SE(svytotal_k_winsorized(~P85, ech.stsi, k = k))
}
plot(totals_k_winso, xlab = "Value of k", ylab = "total")
abline(h = P85_total, col = "red", lwd = 2)
```

### Trimmed

```{r}
svytotal_trimmed(~P85, ech.stsi, LB = 0, UB = 0.90)
```

## Weighted

### Mean

```{r}
weighted_mean_winsorized(stsrswor_data$P85, poids_stsi, LB = 0, UB = 0.90)
```

```{r}
weighted_mean_k_winsorized(stsrswor_data$P85, poids_stsi, k = 2)
```

```{r}
weighted_mean_trimmed(stsrswor_data$P85, poids_stsi, LB = 0, UB = 0.90)
```

### Totals

```{r}
weighted_total_winsorized(stsrswor_data$P85, poids_stsi, LB = 0, UB = 0.90)
```

```{r}
weighted_total_k_winsorized(stsrswor_data$P85, poids_stsi, k = 2)
```

```{r}
weighted_total_trimmed(stsrswor_data$P85, poids_stsi, LB = 0, UB = 0.90)
```

## Stratum jumpers

Create a observation in not the right strata : big cities in small or medium strata
